---
title: "Covid-19 Dashboard"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
    theme: journal
---

```{r setup, include=FALSE}
library(flexdashboard)
```

```{r, include=FALSE}
# EDA AND DATA MANIPULATION

# Loading necessary paclages to R
library(dplyr)

library(tidyverse)

# Loading the Covid 19 global data 
covid_19_data <- read.csv("WHO-COVID-19-global-data.csv")

# Checking for NA values in the entire dataframe
na_summary <- sapply(covid_19_data, function(col) sum(is.na(col)))

# Display the summary of NA values
print(na_summary) # No NA values were found to disrupt data 

# Case Counts
# Selecting the important columns for Covid 19 case count 
covid_19_data_cases <- covid_19_data %>%
  select(Date_reported, Country_code, Country, New_cases, Cumulative_cases)

# Keeping latest data regarding cumulative cases from each Country for the year 2022 (December.31.2022)
covid_19_data_latest_cases <- covid_19_data %>%
  select(Date_reported, Country_code, Country, Cumulative_cases)

covid_19_data_latest_cases <- covid_19_data_latest_cases %>%
  filter(Date_reported == "2022-12-31")

# Death Counts
# Selecting the important columns for Covid 19 death count 
covid_19_data_deaths <- covid_19_data %>%
  select(Date_reported, Country_code, Country, New_deaths, Cumulative_deaths)

# Keeping latest data regarding cumulative deaths from each Country for the year 2022 (December.31.2022)
covid_19_data_latest_deaths <- covid_19_data %>%
  select(Date_reported, Country_code, Country, Cumulative_deaths)

covid_19_data_latest_deaths <- covid_19_data_latest_deaths %>%
  filter(Date_reported == "2022-12-31")

# Loading the necessary packages for API search
library(wbstats)

# Viewing the data set generated from the wbstats package regarding Country information
  # We can view the average income per Country by looking at the "country" column and the "income_level" column.
View(wb_countries())

# Viewing the GDP data for each Country 
df_gdp <- wb_data("NY.GDP.MKTP.CD")

filtered_df_gdp <- df_gdp %>%
  filter(date %in% c(2020, 2021, 2022))

# Viewing the unemployment levels per Country
df_unemployment <- wb_data("SL.UEM.TOTL.ZS")

filtered_df_unemployment <- df_unemployment %>%
  filter(date %in% c(2022))

# Viewing the population data for each Country 
df_population <- wb_data("SP.POP.TOTL")

filtered_df_population <- df_population %>%
  filter(date %in% c(2022))
```

```{r, include=FALSE}

# CALCULATIONS

# Calculating Prevalence (Confirmed cases by Population)
  # Looking at the "covid_19_data_latest_cases" data frame and "filtered_df_population"data frame
prevalence <- covid_19_data_latest_cases %>%
  left_join(filtered_df_population, 
            by = c("Country_code" = "iso2c")) %>%
  select(-iso3c:-date,-unit:-last_updated) %>%
  mutate(prevalence_covid = (Cumulative_cases / SP.POP.TOTL))
View(prevalence)

# Case Fatality Rate (Deaths per Infected Population)
# Selecting the important columns for Covid 19 Population Count 
covid_19_data_cases_fatality <- covid_19_data %>%
  select(Date_reported, Country_code, Country, Cumulative_cases, Cumulative_deaths)

covid_19_data_cases_fatality <- covid_19_data_cases_fatality %>%
  filter(Date_reported == "2022-12-31")

covid_19_data_cases_fatality <- covid_19_data_cases_fatality %>%
  mutate(Case_Fatality_Rate = Cumulative_deaths / Cumulative_cases)

# Mortality Rate (Deaths per Total Population)
# Looking at the "covid_19_data_latest_deaths" data frame and "filtered_df_population"data frame

mortality_rate_per_country <- covid_19_data_latest_deaths %>%
  left_join(filtered_df_population, 
            by = c("Country_code" = "iso2c")) %>%
  select(-iso3c:-date,-unit:-last_updated) %>%
  mutate(Mortality_rate = (Cumulative_deaths / SP.POP.TOTL))

  
```

# Demographics

## Row {data-height="100"}

Covid-19 data from Jan 1, 2020 and Dec 31, 2022 was reported below in each country as reported by the World Health Organization (WHO).
To View Mortality Rate, Case Fatality Rate and Prevalence of COVID-19, please zoom in and click on a country!

Note: Countries in white did not have available COVID-19 Data for the calculations, also some countries may have NA values due to lack of information

### Interactive World Map

```{r}
library(flexdashboard)
library(leaflet)
library(dplyr)
library(rnaturalearth)
library(htmltools)

map_df <- cbind(mortality_rate_per_country$Country, mortality_rate_per_country$Mortality_rate, covid_19_data_cases_fatality$Case_Fatality_Rate, prevalence$prevalence_covid)

# Get world shapefile
world_shape <- ne_countries(returnclass = "sf")

# Merge the shapefile with your data
merged_data <- merge(world_shape, map_df, by.x = "name", by.y = "V1")

leaflet(merged_data) %>%
  addProviderTiles("OpenStreetMap.Mapnik") %>%
  addPolygons(stroke = TRUE,
              fillColor = ~"blue",
              fillOpacity = 0.5,
              color = "white",
              weight = 1,
              popup = ~paste("Country: ", name, "<br>",
               "Mortality Rate: ", V2, "<br>",
               "Case Fatality Rate: ", V3, "<br>",
               "Prevalence: ", V4)

  )

  
```

# Covid + Socioeconomic Factors

## Row {data-height="100"}

Chart 1 - Case Fatality Rate according to GDP: In the graph below, case fatality rates are being compared to each Country’s Gross Domestic Product (GDP) to form a scatter plot. The majority of the data points form a vertical line, which signifies that even with different rates of fatality, GDP stays consistent throughout. There are a few outliers, but the majority do not stray from the line.  

Chart 2 - Case Fatality according to Income: In the second graph, case fatality rates are being compared to the different levels of the average population's income of various countries. Income levels were divided into 5 categories:

- High Income
- Low Income 
- Lower Middle Income 
- Not Classified 
- Upper Middle Income

As can be seen by the box plots, the mean case fatality is higher for the "Low Income" and "Lower Middle Income" categories. This signifies that there are more case fatalities for countries in which the average person has a lower income. 
The mean case fatality is lower for the "Upper Income" and "Upper Middle Income" categories, with the prior have the lowest mean.

Chart 3 - Case Fatality Rate According to Unemployment Rate: In the third graph, case fatality rates are being compared to the unemployment rates of each country. As can be seen by the scatter plot, a positive relationship is established between the two variables. The higher the unemployment rate, the higher the case fatality rate is.

## Row {data-height="100"}
### Chart 1

```{r}
library(ggplot2)
# visualization of case fatality by WHO region

ggplot(covid_19_data, aes(x = WHO_region, y = Cumulative_cases, fill = WHO_region)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Case Fatality by WHO Region",
       x = "WHO Region",
       y = "Case Fatality Rate") +
  theme_minimal() +
  scale_fill_discrete(name = "WHO_region")

```

### Case fatality by GDP

```{r}
# Choose only 2022 data
gdp_22 <- df_gdp %>% 
  filter(date == 2022)
View(gdp_22)
# Rename to same column name is case fatality data frame. This will allow for a merge
gdp_22 <- gdp_22 %>%
  rename(Country = country)

merge_gdp_cfr <- merge(gdp_22, covid_19_data_cases_fatality, by = "Country", all.x = TRUE)

library(ggplot2)
# Create plot
ggplot(merge_gdp_cfr, aes(x= NY.GDP.MKTP.CD, y = log(Case_Fatality_Rate))) +
  geom_point() +
  labs(title = "Case Fatality Rate according to GDP",
       x = "GDP",
       Y = "Case Fatality Rate") + 
  scale_x_continuous(limits = c(min(merge_gdp_cfr$NY.GDP.MKTP.CD), max(merge_gdp_cfr$NY.GDP.MKTP.CD)))

```

## Row {data-height="100"}
### Case Fatality by Income

```{r}

# Only take 2022 data 
gdp_22 <- df_gdp %>% 
  filter(date == 2022)
View(gdp_22)

pop_22 <- df_population %>%
  filter(date == 2022)
View(pop_22)

pop_22 <- pop_22 %>%
  rename(Country_code = iso2c)

merge_pop22_cfr <- merge(pop_22, covid_19_data_cases_fatality, by = "Country_code", all.x = TRUE)

View(merge_pop22_cfr)
# rename column for ease of use 
merge_pop22_cfr <- merge_pop22_cfr %>%
  rename(iso2c = Country_code)
# merge df to one to ensure same number of rows
merge_pop22_cfr_income <- merge(merge_pop22_cfr, wb_countries(), by = "iso2c", all.x = TRUE)

library(ggplot2)
# ensure income level is factor 
merge_pop22_cfr_income$income_level <-as.factor(merge_pop22_cfr_income$income_level)

# Create plot 
ggplot(merge_pop22_cfr_income, aes(x = income_level, y= Case_Fatality_Rate)) +
  geom_boxplot(fill = "lightgrey") + 
  labs(title = "Case Fatality by Income Level",
       x= "Income Level",
       y = "Case Fatality Rate")
  
```

### Case fatality by Unemployment Rate

```{r}
# selct 2022 data
unemployment_22 <- df_unemployment %>% 
  filter(date == 2022)
View(unemployment_22)
# rename column to allow for merge of dataframes 
unemployment_22 <- unemployment_22 %>%
  rename(Country = country)

merge_unemployment_cfr <- merge(unemployment_22, covid_19_data_cases_fatality, by = "Country", all.x = TRUE)

library(ggplot2)
# Create plot
ggplot(merge_unemployment_cfr, aes(x= SL.UEM.TOTL.ZS, y = log(Case_Fatality_Rate))) +
  geom_point() +
  labs(title = "Case Fatality Rate according to Unemployment Rate",
       x = "Unemployment Rate",
       Y = "Case Fatality Rate") + 
  scale_x_continuous(limits = c(min(merge_unemployment_cfr$SL.UEM.TOTL.ZS), max(merge_unemployment_cfr$SL.UEM.TOTL.ZS)))

```

``` {r, include = FALSE}

# ANOVA

library(dplyr)
# Unemployment rate was chosen as the socioeconomic indicator of interest
merge_unemployment_cfr <- merge_unemployment_cfr%>%
mutate(SL.UEM.TOTL.ZS = case_when(
  SL.UEM.TOTL.ZS > 0 & SL.UEM.TOTL.ZS <= 10 ~ 1,
  SL.UEM.TOTL.ZS > 10 & SL.UEM.TOTL.ZS <= 20 ~ 2,
  SL.UEM.TOTL.ZS > 20 & SL.UEM.TOTL.ZS <= 30 ~3,
         TRUE ~ NA_integer_
  ))

levels(merge_unemployment_cfr$SL.UEM.TOTL.ZS)
merge_unemployment_cfr$SL.UEM.TOTL.ZS <- as.factor(merge_unemployment_cfr$SL.UEM.TOTL.ZS)

levels(merge_unemployment_cfr$Country)
merge_unemployment_cfr$Country <- as.factor(merge_unemployment_cfr$Country)

# Change column title 
View(merge_unemployment_cfr)
merge_unemployment_cfr <- merge_unemployment_cfr %>%
  rename(unemploy_rate = SL.UEM.TOTL.ZS)

# Run ANOVA on CFR and unemployment rate 
model <- anova(lm(Case_Fatality_Rate ~ unemploy_rate, data = merge_unemployment_cfr))

p_value <- model$"Pr(>F)"[1]

# Conduct bonferroni post hoc test on p value
adjusted_val <- p.adjust(p_value, method = "bonferroni")
print(adjusted_val)

# No significant findings, adjusted p-value greater than 0.05

```